### 2015-09-02 paper/manuscript.tex
start manuscript. add acknowledgement to takeuki

\documentclass{article}

\begin{document}

\section{Acknowledgmenets}

Takeuki Uno for the insight provided to find the set of all maximum matching algorithms.

\end{document}