### 2015-10-08 code/07_max-match_vs_redundancy.Rmd
alles

---
title: "redundancy vs maximum matchings"
author: "Fernando Cagua"
date: "5 October 2015"
output: html_document
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
library(ggtern)

opts_knit$set (root.dir=normalizePath('../'))
opts_chunk$set (fig.path = "./figures/07/")
opts_chunk$set (cache.path = "./cache/07/")
opts_chunk$set (fig.width = 11, 
                fig.height = 9,
                out.width = "858px")
```

```{r}
# load utility functions
nn <- list.files("./code/functions", full.names = TRUE) %>% lapply(source)
# load netw
net <- list.files("./data/networks/", full.names = TRUE) %>%
    lapply(readRDS)
names(net) <- list.files("./data/networks/", full.names = FALSE) %>%
    stringr::str_split("\\.") %>% lapply(`[`, 1) %>% unlist ()
```

Let's check if the redundancy matches the maximum matching results.

## Read redundancy

```{r}
node_redundancy <- read.csv(file = "./data/node_redundancy.dat")

node_re <- node_redundancy %>% 
	plyr::ddply("net", function(x){
		this_net <- net[[first(x$net)]] %>%
	# select only the largest connected component of the network
			keep_largest_component()
		nodes <- igraph::V(this_net)[x$n]$name
		x %>%
			dplyr::mutate(node = nodes)
	}) %>% dplyr::tbl_df()
```

## Read maximum matchings

```{r}
folder <- "./data/maximum_matchings_summary/" 
files <- list.files(folder)
node_mm <- plyr::adply(files, 1, function(x){
	read.csv(file.path(folder, x)) %>%
		dplyr::mutate(net = unlist(stringr::str_split(x, "_"))[1],
									type = unlist(stringr::str_split(x, "_"))[2], 
									keep = unlist(stringr::str_split(x, "_"))[3],
									batch = unlist(stringr::str_split(x, "_"))[4])
}) %>%
	dplyr::select(-X1) %>%
	dplyr::group_by(net, type, keep, node) %>%
	dplyr::summarise(freq = sum(freq))

```

## Join them

```{r}
dplyr::full_join(node_mm, node_re) %>%
	dplyr::filter(!is.na(freq)) %>%
	dplyr::group_by(net, type, keep) %>%
	dplyr::mutate(r = rank(freq, ties.method = "first"),
								f_f = freq/max(freq)) %>%

	ggplot() +
	geom_point(aes(x = r, y = f_f, fill = robustness), shape = 21) +
	facet_wrap(type ~ net, scales = "free")
```