### 2015-12-05 code/11_compare_n-drivers_random.Rmd
analise networks versus randomisations for n control nodes

---
title: "Comparing n_drivers with random networks"
author: "Fernando Cagua"
date: "3 December 2015"
output: html_document
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
library(ggtern)

opts_knit$set (root.dir=normalizePath('../'))
opts_chunk$set (fig.path = "./figures/10/")
opts_chunk$set (cache.path = "./cache/10/")
opts_chunk$set (fig.width = 11, 
                fig.height = 9,
                out.width = "858px")
```

```{r}
# load utility functions
nn <- list.files("./code/functions", full.names = TRUE) %>% lapply(source)
# load netw
net <- list.files("./data/networks/", full.names = TRUE) %>%
	lapply(readRDS)
names(net) <- list.files("./data/networks/", full.names = FALSE) %>%
	stringr::str_split("\\.") %>% lapply(`[`, 1) %>% unlist ()
```

Get number of controls

```{r}
# generate parameter space
space <- expand.grid(type = c("z-bi", "weight", "AB", "BA"), 
                                         net = 1:length(net)) %>%
    dplyr::inner_join(data.frame(type = c("z-bi", "AB", "BA", "weight", "weight"),
                                                             keep = c("all", "A", "B", "A", "B")))

# select the corresponding one for this computation

driver_nodes <- 1:nrow(space) %>%
	plyr::laply(function(i){
		p <- space[i,]
		n <- net[[p$net]] %>%
			keep_largest_component() %>%
			bipartite_digraph(type = p$type, keep = p$keep)
		igraph::E(n)$weight <- 1
		n %>%
			n_unmatched_vertex()
	})

space %<>%
    dplyr::mutate(n_driver = driver_nodes,
                                type.keep = interaction(type, keep))

net_names <- plyr::ldply(net, function(x){
    length(igraph::V(x))
}) %>%
    dplyr::mutate(net = 1:nrow(.)) %>%
    dplyr::rename(n_species = V1,
                                net_name = `.id`)

n_control <- dplyr::inner_join(space, net_names) %>%
    dplyr::mutate(n_no_driver = n_species - n_driver)
```

Generate a random network and calculate the number of no drivers

```{r}
library(foreach)
library(doMC)
registerDoMC(cores = 4)

# generate parameter space
space <- expand.grid(type = c("z-bi", "weight", "AB", "BA"), 
										 net = 1:length(net), 
										 method = c("r0", "c0", "quasiswap")) %>%
	dplyr::inner_join(data.frame(type = c("z-bi", "AB", "BA", "weight", "weight"),
															 keep = c("all", "A", "B", "A", "B")))

# number of desired permutations
nperm <- 999

random_n <- space %>%
	# for each network and randomisation type
	plyr::ddply(c("net", "method"), function(p){
		nn <- net[[p$net[1]]] %>%
			keep_largest_component() %>%
			bipartite_digraph(type = p$type, keep = p$keep)
		igraph::E(nn)$weight <- 1

		# generate a set of random networks
		random_net <- get_random_net(nn, as.character(p$method[1]), nperm)

		# for each of the networks
		plyr::ldply(random_net, function(x){
			1:nrow(p) %>%
				# for each of the directionality schemes
				plyr::ldply(function(i){
					p <- p[i,]
					# calculate number of control nodes
					x %<>% keep_largest_component() %>%
						bipartite_digraph(p$type, p$keep)
					data.frame(type = p$type, keep = p$keep, 
										 n_driver = x %>%
										 	n_unmatched_vertex())
				})
			# make it pararell
		}, .parallel = T)
	}, .progress = "text")

```

let's have a quick view

```{r}
ggplot(random_n) + 
	geom_boxplot(aes(x = as.character(interaction(type, keep)), y = n_driver, fill = method), colour = "grey10") + 
	geom_point(data = n_control, 
						 aes(x = interaction(type, keep), y = n_driver), colour = "red") +
	facet_wrap(~ net, scales = "free") +
	theme_classic()
```

Save the resulys
```{r}
write.csv(random_n, "./data/random_n_driver.csv", row.names = FALSE)
```

Calculate z scores

```{r}
random_n_z <- random_n %>% 
	dplyr::group_by(net, type, keep, method) %>%
	dplyr::summarise(n_driver_u = mean(n_driver, na.rm = T),
									 n_driver_sd = sd(n_driver, na.rm = T)) %>%
	dplyr::full_join(n_control) %>% 
	dplyr::group_by() %>%
	dplyr::mutate(n_driver_z_score = (n_driver - n_driver_u)/n_driver_sd)

random_n_z %>% 
	ggplot(aes(x  = type, y = n_driver_z_score)) +
	geom_hline(yintercept = c(-2, 2), linetype = 2) +
	geom_boxplot(aes(fill = method)) +
	theme_classic()
```

Diversity ratio and z scores

```{r}
net_info <- read.csv("./data/ntw_info.csv")

net_info %<>%
	dplyr::mutate(ratio_pol_pla = n_pol/n_pla)

random_n_z %>%
	dplyr::full_join(net_info) %>%
	dplyr::filter(method == "c0") %>%
	ggplot(aes(x = n_driver_z_score, y = ratio_pol_pla, colour = as.character(interaction(type, keep)))) +
	geom_point() +
	geom_smooth(method = "lm")

```