### 2017-05-04 paper/drafts_and_revisions/submission_2016Jul11/covering_letter_dbs_2016-07_12.tex
Change folder names

\documentclass[10pt, a4paper]{letter}

\begin{document}

\begin{letter}{
       %\vspace*{20pt}
       Prof. Tim Coulson\\
       Editor in Chief\\
       Ecology Letters}

\opening{Dear Proffesor Coulson,}

We are submitting the manuscript entitled ``Biotic invasions reduce the manageability of pollination networks'' to be considered for publication at \emph{Ecology Letters}.

Central to this manuscript is the introduction, development, and application
of the theoretical concept of \textbf{ecological manageability}. Based on the
patterns of interactions between species, this idea can provide two main
pieces of information. First, it allows us to quantify the \textit{community
manageability}---an indication of the difficulty of controlling the abundances
of all species in the community. Second it allows to identify \textit{driver
species}--- those species with disproportionate ability to alter other
species' abundances. Here, we use this concept as a framework to study the
changes induced by biotic invasions to pollination networks.

Multiple empirical observations have suggested that the structural changes
caused by invasions may lead to increased ecosystem resilience. Our study is
the first to use empirical data to provide strong theoretical support for this
idea. Furthermore, our approach to identify driver species and quantify their
relative importance is unique in that it is based directly on the aspects of
network structure that determine the behavior of the community as a holistic
dynamical system. Our results also highlight the extent to which some
species---invaders in particular---can drive the dynamics of other species
across their community, and also identify ``asymmetric dependence'' as the
potential key mechanism driving the interspecific differences.

Despite our focus, understanding how drivers of ecosystem change affect
ecosystem dynamics is not exclusive to mutualistic communities. Indeed, we
envisage the application of the concepts we introduce here to answer a wide
range of ecological questions in communities structured by different types of
interactions. As such, we have sought to highlight the methodology necessary
for its implementation---an extension of recently developed tools at the
interface of complex systems and control theory---thoroughout the manuscript.

While our study fundamentally lies within ecological theory, we cannot
overlook its relevance to ecological applications. Indeed, we demonstrate in
the paper itself that concepts like ``manageability'' have the potential to
stimulate much needed research that shortens the gap between theory and
management. Hopefully our study and more like it will set us on the track to
informed, rather than hopeful, ecosystem interventions.

Lastly, please note that the enclosed work represents a novel contribution for
all co-authors involved and no related work published, in press, or submitted
during this or last year has been cited therein.

\closing{Regards,}

Daniel B. Stouffer

\end{letter}

\end{document}