### 2018-03-27 publication/manuscript.Rmd
Start trying to make a new intro

---
title: "Network control"
classoption: a4paper
csl: journal-of-ecology.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{caption}
- \usepackage{lineno}
documentclass: artikel1
bibliography: references.bib
---

\doublespacing
\linenumbers

**E. Fernando Cagua^1^, Kate L. Wootton^1,2^, Daniel B. Stouffer^1^**

^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Current address: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, SE-750 07 Uppsala, Sweden

**Author for correspondence:** Daniel Stouffer (daniel.stouffer@canterbury.ac.nz) - +64 3 364 2729 - Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8140, New Zealand

\newpage

# Introduction

Networks have used to study the role of species in their community. Most approaches are rooted on the assumption that there is a link between a species centrality and its importance—its ability to modify the abundance of other species. However, recent studies show that definitions of role based on centrality, although simple, might not be a good proxy for the importance of a species for ecosystem functioning. Here we show that using a control theory can yield novel insights about a species role in the community.

```{r example_network, include = F}
loadd(example_net)

base_net <- example_net
	bi_net <- example_net %>%
		keep_largest_component() %>%
		bipartite_digraph(type = "z-bi", weight.type = "asymetry") 
	as_net <- example_net %>%
		keep_largest_component() %>%
		bipartite_digraph(type = "weight", weight.type = "asymmetry")

	fam <- "Helvetica"

	# layout is based on the bidirectional network
	set.seed(1)
	bi_net_uw <- bi_net
	E(bi_net_uw)$weight <- 1
	lay <- bi_net_uw %>%
		layout_nicely()
	v_c <- "white"
	l_c <- "black"
	l_s <- 0.9
	ma <- c(0,0,0,0)
	label_pos_x <- -1
	label_pos_y <- 1
```