### 2018-04-06 publication/manuscript.Rmd
Finish rewrite of the intro

---
title: "Ecological network control"
classoption: a4paper
csl: journal-of-ecology.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{caption}
- \usepackage{lineno}
documentclass: artikel1
bibliography: references.bib
---

\doublespacing
\linenumbers

**E. Fernando Cagua^1^, Kate L. Wootton^1,2^, Daniel B. Stouffer^1^**

^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Current address: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, SE-750 07 Uppsala, Sweden

**Author for correspondence:** Daniel Stouffer (daniel.stouffer@canterbury.ac.nz) - +64 3 364 2729 - Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8140, New Zealand

\newpage

# Introduction

Within community ecology, a complex-systems approach has led to the development of analytical and simulation tools with which to understand the role of species embedded in a network of interactions [@Bascompte2009; @Stouffer2012]. 
A particularly relevant dimension of a species role for ecosystem functioning is that which relates to its ability to alter the abundance of other species in the community. 
This ability is often referred as species importance or "keystoness". 
A significant proportion of the network tools used to estimate a species ability to modify its community is based on the assumption that there is a link between a species centrality and its importance. 
However, recent studies have shown that definitions of role based on centrality, although straightforward, might not adequately estimate the relative influence of a species on its community. 
To overcome this issue, we require a framework that is equipped to account for the structure of the interactions in the community, but that also acknowledges the dynamics that determine species abundance and ultimately the state of the community.

Among the various possibilities, *control theory* appears to be a strong candidate [@Isbell2013]. 
Widely used in engineering to determine and supervise the behaviour of dynamical systems [@Motter2015], it is well equipped to deal with the many feedbacks present in ecological communities [@Liu2016]. 
Furthermore, advances in this field have established a link between the structure of the network and the way nodes affect each other that is based on well-established advances in both dynamical and complex systems theory. 
As such, a priori assumptions about how particular node metrics, like centrality, are related to network functioning are not required
Despite the coincidences between ecological and engineering systems, the concepts and approaches of control theory are not commonly applied in an ecological context. 
This is, in part, because in contrast with ecology, research on the control of complex networks is generally more concerned with driving the state the system as a whole rather than the nodes that compose the network. 
At its fundamental level, control theory is concerned with determining whether a system is at all controllable, this is, whether it can be driven to a desired state or not. 
Our first contribution is to expand existing methods control theory to effectively link the *controllability* of a network with the role of particular nodes.

After overcoming this first hurdle, we then apply the methods to an ecological problem and show how it can be used to gain insight into the role of species in an ecological network. 
Specifically, we outline the approach using a set of ten pairs of uninvaded and invaded plant-pollinator communities. 
We use invaded communities because there is strong empirical evidence showing that invasive species play an important role shaping the abundances of other species. Certainly so in the chosen networks [@Lopezaraiza-Mikel2007; @Bartomeus2008]. 
Hence, this choice offers an opportunity to contrast theoretical observations with empirical evidence. 
First, grounded in the difficulties usually involved with invasive-species eradication and ecosystem restoration [@Woodford2016], we ask whether invaded networks require a greater proportion of species to be managed to achieve the same level of control. 
Second, we ask—from a control theoretic perspective—wether there are differences among species in the role they play at driving the population of other members of the community and the factors related to these differences. 
Finally, we compare the proposed approach to current methods based on species centrality and show how these methods are indeed valuable but only paint a limited picture in regards to the "keystoness" of a species.

```{r example_network, include = F}
loadd(example_net)

base_net <- example_net
	bi_net <- example_net %>%
		keep_largest_component() %>%
		bipartite_digraph(type = "z-bi", weight.type = "asymetry") 
	as_net <- example_net %>%
		keep_largest_component() %>%
		bipartite_digraph(type = "weight", weight.type = "asymmetry")

	fam <- "Helvetica"

	# layout is based on the bidirectional network
	set.seed(1)
	bi_net_uw <- bi_net
	E(bi_net_uw)$weight <- 1
	lay <- bi_net_uw %>%
		layout_nicely()
	v_c <- "white"
	l_c <- "black"
	l_s <- 0.9
	ma <- c(0,0,0,0)
	label_pos_x <- -1
	label_pos_y <- 1
```

# References