### 2018-05-25 paper/cover-letter.tex
Add cover letter (tex file)
Addresses #8

\documentclass{letter}
\usepackage[a4paper, total={5in, 9.1in}]{geometry}
\usepackage{setspace}
\signature{Fernando Cagua \\ (in behalf of all authors)}
\address{Centre for Integrative Ecology \\ School of Biological Sciences \\ University of Canterbury \\ Private Bag 4800 \\ 8140 Christchurch \\ New Zealand}
\begin{document}
\begin{letter}{James Ross \\ Assistant Editor \\ Journal of Ecology}
\opening{Dear James:}

\onehalfspacing

We are thankful to the senior and associate editors, Prof. James Ross and Prof. Nicole Rafferty, and two anonymus reviewers for the insightful comments to improve the manuscript. 
We are also grateful for the opportunity to resubmit. 
In the attached documment you can find a detailed response to each of their comments. 
Here we summarise the main changes to the manuscript.

Briefly, our paper explores the possible applications of control theory in an ecological context. 
We see potential in our paper to contribute to both applied and theoretical aspects in ecology. 
However in our initial submission we lingered midway between these two directions and—as the editors and reviewers pointed out—we failed to address either option successfuly.
We have re-inspected our manuscript and tink that the theoretical insights are stronger. 
As a consequence in this revision we have chosen to focus on the conceptual aspect of our approach at the expense of numerous, but weak, connections with conservation and applied ecology.

Adopting a primarely conceptual focus allowed us to be clearer about the justification, methodology and potential insights of our approach. 
However, improving clarity while successfully tying the conceptual motivation with our findings required more than some changes to some paragraphs or sections of the manuscript. 
Therefore, we have rewriten almost entirely the manuscript and the supplementary materials. 
We maintain the methodology and findings, but there has been a major change in the way we justify the study, introduce novel concepts to the ecological domain, and present our conclusions.

We look forward to further feedback and hope that the editors and reviewers find these changes satisfactory.

\closing{Regards,}
\end{letter}
\end{document}