### 2018-06-08 paper/supp-info.Rmd
Add results of controllability models
Close #13

---
title: "Ecological network control"
subtitle: "Supplementary Information"
author: "Fernando Cagua, Kate L. Wootton, Daniel B. Stouffer"
classoption: a4paper
csl: journal-of-ecology.csl
documentclass: artikel1
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{xr}
  - \externaldocument[M-]{manuscript}
bibliography: references.bib
---

\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}

\renewcommand\thetable{S\arabic{table}}    
\setcounter{table}{0}

\renewcommand{\thesection}{S\arabic{section}}
\setcounter{section}{0}

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
```

# Characteristics of the empirical networks {#empirical-networks}

```{r}
drake::loadd(metadata)
drake::loadd(network_properties)

net_prop <- metadata %>%
	dplyr::mutate(n_s_t = n_pla + n_pol) %>%
	dplyr::select(-n_pla, -n_pol) %>%
	dplyr::inner_join(network_properties, by = "net_name")

```

The networks studied had species richness ranging between `r min(net_prop$n_s_t)` and `r max(net_prop$n_s_t)` when considering only the largest component in each network). 
As shown by the network asymmetry *AS* [@bluthgen_specialization_2007], the networks had a low ratio of plants to pollinators overall. 
Furthermore, the networks had relatively low levels of nestedness [when measured using the quantitative version of the NODF index; @almeida-neto_straightforward_2011]. Details for each network can be found in the Table \@ref(tab:table-network-properties).

```{r table-network-properties, results = "asis", sanitize = FALSE}
net_prop %>%
	dplyr::select(-net_name, -count, -net) %>%
	dplyr::filter(study != "ballantyne") %>%
	dplyr::arrange(study, site, inv) %>% 
	dplyr::mutate(invader = dplyr::case_when(
				 	invader == "car" ~ "\\textit{Carpobrotus affine}",
				 	invader == "op" ~ "\\textit{Opuntia stricta}",
				 	invader == "imp" ~ "\\textit{Impatients grandulifera}",
				 	TRUE ~ "—"
				 ),
				 invader = dplyr::if_else(inv, invader, "—")) %>%
	dplyr::mutate(site = rep(1:n_distinct(site), each = 2),
				 inv = dplyr::if_else(inv, "yes", "no"),
				 study = dplyr::case_when(
				 	study == "bartomeus" ~ "Cap de Creus, Spain",
				 	TRUE ~ "Bristol, United Kingdom"
				 )) %>% 
  dplyr::select(site, invader, n_sp, n_pla, n_pol, connectance, web.asymmetry, weighted.NODF, study) %>%
  dplyr::rename_(
    # "$R$" = "n_s_t",
					"$n_p$" = "n_pla",
					"$n_a$" = "n_pol",
					"location" = "study",
					"$n_s$" = "n_sp",
					"$c$"= "connectance", 
					"NODF" = "weighted.NODF",
					# "$\\bar{a}$" = "interaction.strength.asymmetry",
					"$AS$" = "web.asymmetry") %>% 
  knitr::kable(escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c("", "\\addlinespace"),
               caption = "Properties of the analysed plant-pollinator communities. Here show the number of species ($n_s$), the number of plants ($n_p$), the number of pollinators ($n_a$), the network connectance ($c$), the network asymmetry ($AS$), and the network nestedness (NODF index). All properties correspond to the network's largest component. British networks were assembled by Lopezaraiza-Mikel et al. (2007), Spanish were networks assembled by Bartomeus et al. (2008).")
```

# Structural controllability {#structural-controllability}

Structural controllability avoids the limitation of not knowing the exact values of the matrices *A* and *B*. 
Using structural controllability can be boiled down to two conditions: a system is controllable if there are no inaccessible nodes or dilations. A node is inaccessible if there are no directed paths between it and the input nodes. "Dilations are subgraphs in which a small subset of nodes attempts to rule a larger subset of nodes. In other words, there are more 'subordinates' than 'superiors'" [@liu_control_2016].

The goal of structural controllability is to use the information contained in **A** to generate a supportable estimate of **B**.
This focus allows us to gain insight of the inherent controllability of a network, and the roles of the species that compose it, without being overly dependent on the particular choices of how the system dynamics are modelled or characterised. 
The trade-off of this approach is that, because of the assumption of linearity, structural controllability alone does not allow us to fully design the time-varying control signal *u(t)* that can drive the system from one particular equilibrium to another. 
Nevertheless, the lessons gained when assuming linearity---at both the network and the species level---are a prerequisite for eventually understanding nonlinear control [@liu_controllability_2011; @liu_control_2016].

# Maximum matching

## Finding a single maximum matching {#one-maximum-matching}

Our approach to finding the minimum number of driver nodes relies on finding maximum matchings. 
We start with a directed network in which the direction of the link represents the direction of control (Figure \@ref(fig:fig-maximum-matching-procedure) left panel). 
We then construct an alternative representation of the directed network in which each node of the directed network is represented by two nodes that indicate their outgoing and incoming links respectively (Figure \@ref(fig:fig-maximum-matching-procedure) centre panel). 
Finding a maximum matching in this alternative representation is equivalent to finding the largest possible set of edges in which one node on the left-hand side is connected to at most one node on the right-hand side. 
To find the maximum matching we use the push-relabel algorithm implemented in `max_bipartite_matching` in the R package igraph 1.0.1 [@Csardi2006a]. 
Once we have the matching (shown in the Figure \@ref(fig:fig-maximum-matching-procedure) center panel) it is then easy to identify the roles of each node in this representation: nodes on the top-level that are connected to a matched link (dark purple) are superior while those connected to a matched link on the bottom-level are matched. 
This information can then be mapped back to the original representation to identify the control paths and the driver nodes in the network (Figure \@ref(fig:fig-maximum-matching-procedure) right panel).

```{r}
drake::loadd(fig_supp_matching)
```

```{r fig-maximum-matching-procedure, fig.height=fig_supp_matching$height, fig.width=fig_supp_matching$width,fig.cap = "Finding a maximum matching in a complex network. (left) Directed network that indicate the direction of control between species. (center) Alternative bipartite representations of the directed networks. (right) The matchings in the bipartite representation mapped back to the original network."}
replayPlot(fig_supp_matching$plot)
```

To further illustrate our methodology here, we also show the approach for the smallest of our empirical networks, the uninvaded network at site 10 (Table \@ref(tab:table-network-properties); Figure \@ref(fig:fig-small-network)). 
This network is composed of 19 species of which three are non-invasive plants and the other 16 are pollinators. 
The one-to-one relationship between matched and superior nodes implies that in order to achieve full network controllability, most pollinators would be unmatched, and hence are classified as driver nodes that require external intervention. 
At the same time, both plants in the community, *Heracleum sphondylum* and *Rubus fructicosus*, and one of the pollinators, *Orthotylus/Lygocorus*, tend to be classified as superior nodes.

```{r}
drake::loadd(fig_small_network)
```

```{r fig-small-network, fig.height=fig_small_network$height, fig.width=fig_small_network$width,fig.cap = "Illustration of the procedure with an empirical community. The visitation network (top), the number of visits between species pairs is shown on each link. The directed network in which the direction of control is determined based on the mutual dependences (middle). One of the possible maximum matchings of this network (bottom)."}
replayPlot(fig_small_network$plot)
```

## Finding all possible maximum matchings {#all-maximum-matching}

The algorithm implemented in `max_bipartite_matching`, however, is only able to find **one** of possibly many maximum matchings in a network. 
Though one maximum matching is enough to calculate *n~D~* and hence to provide an indication of the manageability of a community, it is not sufficient to estimate the role of individual species. 
To do that, we need to calculate all possible maximum matchings (or, equivalently, all maximal cardinality matchings in weighted networks like ours). 
To do this, we again start from the alternative bipartite representation in Figure \@ref(fig:fig-maximum-matching-procedure)b and assign an identity to each of the links in the network.
We will call this bipartite representation *P*. 
We then construct the line graph of the alternative bipartite representation *L(P)* (Figure \@ref(fig:fig-all-maximum-matching)). 
Each node in *L(P)* represents a link in *P* and these are connected to each other if and only if they share a common node in *P*. 
We then calculate *H*, the complement graph of *L(P)* and identify all of its maximal cliques (Figure \@ref(fig:fig-all-maximum-matching)). 
Here some extra definitions are necessary. 
First, *H* is a graph with the same nodes as *L(P)* but that has a link between two nodes if and only if there is not a link in *L(P)*. 
Second, a clique is a subset of nodes such that all pairs of them are linked. 
Lastly, a maximal clique is a clique such that there are no cliques composed of more nodes [@Gutin2013]. 
In this example, there are two maximal cliques: the one formed by 1, 3 and 5, and the one formed by 2, 3 and 5. 
The final step is then to map these cliques onto the original network to obtain all possible maximal cardinality matchings as shown in Figure \@ref(M-fig:fig-control-configuration) in the main text.

```{r}
drake::loadd(fig_all_matching)
```

```{r fig-all-maximum-matching, fig.height=fig_all_matching$height, fig.width=fig_all_matching$width,fig.cap = "Finding all possible maximum matchings. From left to right: alternative bipartite representation of the directed network in Figure \\@ref(fig:fig-maximum-matching-procedure)a.  Line graph of the network. The complement of the network. The two maximal cliques are shown in dark purple."}
replayPlot(fig_all_matching$plot)
```

## Control adjacency and the input graph {#input-graph}

The algorithm proposed by @zhang_input_2016 enable us to find all minimum driver node sets without calculating all possible maximum matchings. 
The algorithm is based on the construction of the input graph of the directed network. 
Components in the input graph reveal the correlations between nodes from a structural control perspective (Figure \@ref(fig:fig-input-graph)). 
The input graph is constructed based on the *control adjacency* of nodes in a maximum matching. 
Two nodes $x_i$ and $x_j$ are said to be control adjacent if there is a node $x_k$ connecting $x_i$ and $x_j$ with an unmatched link $x_k-x_i$ and a matched link $x_k-x_j$. 
For example $a_1$ is control adjacent to $a_2$ in the top panel of Figure \@ref(fig:fig-input-graph). 
If a node $x_i$ is part of a minimum driver node set then every node in the same component of $x_i$ in the input graph is a possible driver node.

```{r}
drake::loadd(fig_input_graph)
```

```{r fig-input-graph, fig.height=fig_input_graph$height, fig.width=fig_input_graph$width,fig.cap = "Input graph. Two nodes $x_i$ and $x_j$ are said to be control adjacent if there is a node $x_k$ connecting $x_i$ and $x_j$ with an unmatched link $x_k-x_i$ and a matched link $x_k-x_j$. For example, in the top panel, $a_1$ is control adjacent to $a_2$ (via $p_2$), $p_1$ is control adjacent to $p_3$ (via $a_1$) and $p_2$ has no control adjacent nodes. As $a_1$ and $p_2$ belong to the minimum driver node set under the control configuration found with the maximum matching shown, then every node in the same component as these two nodes is also a possible driver node. By combining the nodes from the components with minimum driver node sets is possible to enumerate the different control configurations. For example the network in the top panel has two possible minimum driver node sets $[p_2, a_1]$ and $[p_2, a_2]$ (Figure \\@ref(M-fig:fig-control-configuration)). The network in the middle panel has three minimum driver node sets $[x_1, x_2]$, $[x_1, x_3]$, and $[x_1, x_4]$. Finally the network in the bottom pannel has only one $[x_1]$."}
replayPlot(fig_input_graph$plot)
```

In order to calculate a node's control capacity $\phi$, we harness the fact that nodes can be classified in three groups according to their role in the input graph. 
First,  if $x_i$ belongs to the minimum driver node set and has no adjacent nodes, then it cannot be replaced and will be part of every possible control configuration. 
Second, if $x_i$ does not belong to the minimum driver node set and is not adjacent to any possible input node, then it does not part of the minimum driver node set under any control configuration. 
Finally, if a node $x_i$ belongs to the minimum driver node set and is adjacent to another node $x_j$ in the input graph, then $x_j$ must belong to another minimum driver node to which $x_i$ doesn not belong, in other words $x_i$ and $x_j$ are substitutable.

The frist and the second group have a control capacity of $\phi = 1$ and $\phi = 0$ respectively. 
Meanwhile, the third group has intermediate values of control capacity which can be calculated by computing the number of possible susbstitutions of a node by its control adjacent nodes. 
As a rule of thumb the largest the number of possible substitutions an input node has, the smaller its control capacity.

```{r control-capacity-correlations}
drake::loadd(control_capacity_correlations, this_method)

cc_corr <- control_capacity_correlations %>%
  dplyr::inner_join(metadata, by = "net_name") %>%
  dplyr::filter(study %in% c("bartomeus", "lopezaraiza")) %>%
  dplyr::filter(!is.na(corr), method == this_method) %$%
  corr
```

For larger networks, calculating the control capacities using the input graph is several orders of magnitude faster than computing all possible maximum matchings. 
While the the computation time of the input graph grows linearly with network size, the compuation of all maximum matchings grows polynomially. 
However, the control capacities obtained using the input graph are not identical to those obtained by computing all possible maximum matchings. 
The reason of this discrepancy is that two maximum matchings might result in the same minimum driver node set. 
Nevertheless, the results obtained using both approaches are extremely similar. 
We compared the control capacity obtained by calculating all maximum matchings and that obtained using the input graph in `r length(cc_corr)` of our networks (those for which we were able to calculate all maximum matchings within reasonable time). 
The spearman correlation was very high in all cases, ranging between `r round(min(cc_corr),2)` and `r round(max(cc_corr), 2)` (median `r round(median(cc_corr), 2)`).

## Networks with reciprocal links {#reciprocal-links}

```{r}
drake::loadd(directed_networks, metadata)
recipr <- directed_networks %>%
  purrr::map(igraph::reciprocity, mode = "ratio") %>%
  unlist() %>% {
    dplyr::data_frame(net_name = names(.), prop_bid = .)
  } %>%
  filter_networks_df(metadata)
```

Our procedures to find minimum driver node sets are only adequate when there are no reciprocal (bidirectional) links in the network. 
The proportion of reciprocal links in relatively low. 
In fact the proportion of species pairs that are reciprocally connected in our empirical networks is just `r round(mean(recipr$prop_bid)*100, 2)`%. 
However, reciprocal links are not a rare occurrence as `r sum(recipr$prop_bid != 0)` out of `r nrow(recipr)` empirical networks had at least one. 
To calculate our controllability metrics at both the network (the relative size of the minimum driver node set, $n_D$) and species level (control capacity, $\phi$, and the likelihood of being a superior node, $\sigma$) we first need to generate all possible versions of the of the network that include only non-reciprocal links (Figure \@ref(fig:fig-bidirectional)).

All of these non-reciprocal networks are as likely to be one that best represent the superior/matched relationships in the network. 
As such, we average the three metrics ($n_D$, $\phi$, and $\sigma$) across all the networks without reciprocal links without weighting by the number of possible control configurations.

```{r}
drake::loadd(fig_bidirectional)
```

```{r fig-bidirectional, fig.height=fig_bidirectional$height, fig.width=fig_bidirectional$width,fig.cap = "In order to calculate the controllability metrics for a network with reciprocal links (left), we first decompose the network into several versions that contain no reciprocal links (right). As each of these networks represent a different equiprobable scenario, the controllability metrics for the network with reciprocal links are the unweighted mean of these metrics across the non-reciprocal networks."}
replayPlot(fig_bidirectional$plot)
```

# Empirical vs. random expectation of controllability {#randomisations}

# Controllability models {#controllability-models}

The models were

```{r}
drake::loadd(controllability_models)

replace_terms <- function(x){
  x %>% 
    stringr::str_replace("inv", "inv") %>%
    stringr::str_replace("connectance", "C") %>%
    stringr::str_replace("weighted_NODF", "NODF") %>%
    stringr::str_replace("n_sp", "n_s") %>%
    stringr::str_replace("web_asymmetry", "AS_n") %>%
    stringr::str_replace("interaction_strength_asymmetry", "AS_s") %>%
    stringr::str_replace(stringr::fixed("~"), "\\sim ")
}

AICcmodavg::aictab(controllability_models) %>% 
  dplyr::mutate(Modnames = replace_terms(Modnames)) %>% 
  dplyr::select(Modnames, AICc, Delta_AICc, Cum.Wt) %>%
  dplyr::filter(Delta_AICc <= 8) %>%
  dplyr::rename(model = Modnames, `$\\Delta AICc$` = Delta_AICc, `$AICc$` = AICc, `Cum. Weight` = Cum.Wt) %>%
  dplyr::mutate(model = paste0("$", model, "$")) %>% 
  knitr::kable(escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c(""),
               caption = "Model selection table based on the Akaike Information Criterion for small samples for candidate models of size of the minimum driver node set ($n_D$) in the empirical networks. There was no evidence for any of the evaluated predictors: the invasion status (inv), the network connectance (C), the network nestedness (NODF) the number of species ($n_S$) the network asymmetry ($AS_s$) or the interaction strength asymmetry ($AS_s$). Only candidate models with $\\Delta AIC < 8$ are shown.")
```

# References