### 2019-01-13 paper/reviewers-response.Rmd
Remove reviewers response first round of reviews

---
title: "Keystoneness, centrality, and the structural controllability of ecological networks"
subtitle: "JEcol-2018-0545 - Detailed response to comments"
author: "Fernando Cagua, Kate L. Wootton, Daniel B. Stouffer"
classoption: a4paper
csl: journal-of-ecology.csl
documentclass: artikel1
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
    number_sections: false
header-includes:
  - \usepackage{setspace}
  - \usepackage{xr}
  - \usepackage{lineno}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[M-]{manuscript}
  - \externaldocument[S-]{supp-info}
  - \usepackage[most]{tcolorbox}
  - \definecolor{block-gray}{gray}{0.9}
  - \newtcolorbox{myquote}{colback=block-gray,grow to right by=-10mm,grow to left by=-10mm, boxrule=0pt,boxsep=0pt,breakable}
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
```

We are thankful to the assistant editor James Ross, the senior editor and David Gibson, and three anonymous reviewers for the insightful comments to improve the manuscript. 
In the following sections, you can find a detailed response to each comment raised during the previous round of review.

# Associate Editor's comments:

> "I think this ms now makes much clearer the potential insights to be gained by applying control theory to interaction networks. In particular, the authors have done a nice job of explaining how their findings reinforce existing ideas in the literature on the stability of ecological networks while also acknowledging limitations. The reviewers identify a few remaining issues that should be straightforward to address. I also noticed a few issues"

Thank you for the positive feedback. We have addressed all of the specific issues raised and highlight these in red text in our resubmitted manuscript.

> "*L9*: I think here and throughout much of the text, species should be possessive when followed by a noun ("species role" should be "species' role"). The possessive is currently used in some instances but is inconsistently applied throughout the ms."

*Accepted and changed*. We have checked through the manuscript and corrected this where relevant. Besides *specie's role*, we also changes instances of *species' contribution*, *species' degree*, *species' control capacity*, among others.

> "*L25*: 'every invasive species in our dataset was part of it' Here it was unclear what 'it' refers to (the set of critical species), and the reference to a singular dataset when you have 20 networks in total was a bit confusing to me.

*Accepted and changed*. We have made this clearer as follows: "and included every invasive species in our dataset" (\lr{M-R2-our-dataset})

> "*L23-26*: in these lines, you refer to critical species, but I think the distinction between having high control capacity versus being critical needs to be defined (briefly) earlier (it is stated in *L18-19* that you will quantify both control capacity and identify critical species, but it is not explained how they are distinct)."

*Accepted and changed*. We now identify the distinction between high control capacity and being critical on \lr{M-R2-define-control-capacity}. The text now reads "It also allows us to quantify a species control capacity-its relative importance-and identify the set of species that, collectively, are critical in this context because they have largest possible control capacity."

> "*L27*: unclear what 'its' refers  to."

*Accepted and changed*. We now explicitly state "species' degree" in  \lr{M-R2-species-degree}

> "*L234, 235*: awkward to refer to species as 'they' and 'their'"

*Accepted and changed*. We have changed this to "it" and "its"

> "*L296*: 'More details in Supporting Information.' is a sentence fragment."

*Accepted and changed*. We have corrected this on \lr{M-R2-sentence-fragment}. The text now reads "See Supporting Information Section \@ref(S-visitation-as-proxy) for more details".

> "*L343*: I am not sure about citing Table 2 for the first time in the Discussion. Might it be more useful to cite this in the Introduction?""

We now introduce the glossary at the end of the introdyction on \lr{M-R2-ref-glossary}.

> "*L111, 114*: 'are comprised of' should be 'comprised'. *L339, 432*: 'at' should be 'in'. *L341*: 'allow' should be 'allowed'. *L355*: missing period."

*Accepted and changed*. Thank you for pointing out these mistakes, we corrected all of them.

# Reviewer comments #1

> "I'd like to commend the authors to this revised version of the manuscript on 'Keystoneness, centrality, and the structural controllability of ecological networks'. I thoroughly enjoyed the re-written manuscript, which is now considerably more accessible and clearly targets the community of network ecologists, while not ignoring the importance of the applied relevance of the findings. Overall, I don't have any major comments. I guess the language is in parts on the casual side, yet without compromising the preciseness or clarity of the text. I noticed a few typos or style errors that the authors may want to address. Congratulations on this thought-provoking methodological synthesis."

We are pleased that the reviewer enjoyed the re-written manuscript. We did too!. We also grateful for the typos and style errors identifyed, we have striven to correct them all.

> "A few references require editing to journal style (e.g. remove initials for many citations)"

*Accepted and changed*. We have corrected all citations. One of them, however, (\lr{R2-citation-with-name}) still shows an author's first name initial, as it is necessary to distinguish two authors with the same last name.

> "*L388*: remove repeated 'other'. I suggest that the authors re-write this sentence to improve clarity."

*Accepted and changed*. We have removed the second 'other'. As suggested, we have rewritten the sentence to improve clarity (\lr{M-R2-structural-controllability}). It now reads as follows: "First, structural controllability emphasizes that the effect a species has on the abundance of other species is not independent of the effects of those other species in their community".

> "Reference list needs updating, e.g. Ne'eman et al. is not 2009 but 2010."

*Accepted and changed*. Thanks for pointing this error out. We've corrected the reference.

> "*L314*: open bracket missing. *L358*: change 'instead' to 'Instead'. *L381*: remove second comma. *Fig. S6* header of plot b). correct spelling of 'critical'."

*Accepted and changed*. Thank you for pointing this out, made all these changes as indicated.

# Reviewer comments #2

> "The methods are mostly clear but there is no description of how the 'likelihood of being a superior node' is calculated."

*Accepted and changed*. Thanks for pointing this out. We have modified the sentence mentioning 'the likelihood of being a superior node' in the first paragraph of the secition \@ref(S-reciprocal-links) to "...and wether it is a superior node..." to avoid confusion. Then we define the 'probability of being a superior node' at the end of the section: "It is important to note that a species that is found to be a superior node under a particular control configuration, will be so in all possible control configurations for the particular non-reciprocal network. 
We, therefore denote the average of $\sigma$ across all non-reciprocal versions as the probability of the species being a superior node in its community".

> "*L1364*: 'dynamic persistence'"

*Accepted and changed*: Thank you. To avoid confussion we have removed the word 'dynamic' and now the sentence reads "...our results suggest that structural controllability is closely related to the persistence..." (\lr{M-R2-dynamic-persistence}). It conserves its meaning without loosing much precision.

> "*L1370*: 'dynamic stability'"

*Accepted and changed*: Contrastingly, in this case the word dynanamic distinguishes a particular facet of stability. We mean the ability of the system to return to an equilibrium state after a perturbation. We now include this definition on \lr{M-R2-dynamic-stability} to make the concept clearer.

> I think the concepts presented in this manuscript are a useful addition to the literature on management of ecological networks.  I agree with the authors that it is important that the discussion surrounding what constitutes a "keystone" species should go beyond centrality metrics.

> I found the results and concepts to be generally well presented, particularly graphically, and the authors present a thorough sensitivity analysis to various assumptions.  The methods are mostly clear but there is no description of how the "likelihood of being a superior node " is calculated.

> I would suggest including clarification on the limitations of using control capacity to summarise species' importance, and examples to help the reader build an intuition regarding the control capacity metric.  As pointed out, networks that give rise to a species having a high control capacity can be quite distinct to networks that give rise to a species having a high value of particular centrality metrics.  Providing an example in the Discussion of, for instance, those species that have a control capacity of 1 but a low degree would be very useful.

> "*Discussion*: 'species' strength'

*Accepted and changed*: In the Methods section we deine species' strength as the sum of a species' visits (\lr{M-R2-species-strength-a}). For brevity, we did not define it again in the Disussion, but now we have improved the text and briefly describe it when we mention it there too. The text now reads "When controlling for a species' visitation strength (the sum of a species' visits), which is indirectly a proxy of its abundance..." (\lr{M-R2-species-strength-b}).

> * "desired control trajectory " - does this mean "final state"?

> "'driver node density' (supp p12) -> 'size of the minimum driver node set'?"

> * "dynamic persistence"  (p17,l364)

> "*Figure S1b* is referred to (supp. p4 ) but doesn't exist in the figure"

*Accepted and changed*: Thanks for noticing this issue. We now refer to the figure as "Figure S1 (center panel)".

> "*References*: several references use the term 'plantPollinator', or '...'"

*Accepted and changed*: We have corrected the refeerence that included 'plantPollinator' in the title. We have left the references that include an ellipsis in the author list as they are, as the Journal of Ecology Author Guidelines indicate that this is the preffered format when the reference has more than seven authors.

> * "inherently different " (p18, l 406). The sentence spanning lines 407-410 is quite confusing  - please clarify.

We rephrased this line as follows: "We found that it is not that invasive plants have some different mechanism for influencing the community compared to their native counterparts" \lr{R2-inherrently-different}.

> * "marginally affected by the topology of their interactions"  (p19, l 430)

We have rephrased this line as: "By accounting for network dynamics (even if in a simple way), structural stability incorporates more ecological realism, especially in the extreme scenario in which the structure of interactions within the community only marginally affects the community's state. Both native species and invasive plants tended to attain a high control capacity if they were important to network persistence, were abundant, and depended little on other species." \lr{R2-marginally-affected}.

> "In the species level glm, it was not entirely clear if models were fit to each network individually or if all species were lumped together (the former is what I think was done based on my reading of the text). Regardless I wonder if it would change the results if a mixed effects model was fit to lumped data, using network as a random effect."

*Accepted and changed*: We are sorry for the confusion. In the species level GLM, we lumped all species from different networks together. We have modified the text in \lr{M-R2-one-lumped-model-a} and \lr{M-R2-one-lumped-model-b} to make this clearer. We actually did test for random effects in the model. We tested random effects that allowed for a random intercept for the network (but also its site, and the study it comes from). In all cases, the variance among groups was zero. This means that likely there are no differences on the distribution of control capacity between networks (or site and study) and therefore it was not necessary to include random effects in the models presented in the paper. We have corrected this omission and describe the random effects in \lr{M-R2-random-effects}.

> "How were the model sets derived? Were all possible sub-models compared or was a step-wise approach used?"

*Accepted and changed*: We actually compared all possible candidate models. We have edited the manuscript on \lr{M-R2-all-candidates} to include this information.

> "Reciprocal links were included in the directed graph when dij = dji. Would it have any effect if there was a little wiggle room there, e.g., when the two are close but not equal?"